---
title: "Challenge 5"
subtitle: "She sells bash shells by the seashore"
output:
  html_document:
    df_print: paged
---

# Challenge 5
These questions are based on the following SWC lessions.  You should be able to find answers to very similar problems within these lessons.  All of your answers should use *bash*!  Do not use *R* OR the RStudio *Files* pane.

- [Introducing the Shell](http://swcarpentry.github.io/shell-novice/01-intro/index.html)
- [Navigating Files and Directories](http://swcarpentry.github.io/shell-novice/02-filedir/index.html)
- [Working With Files and Directories](http://swcarpentry.github.io/shell-novice/03-create/index.html)
- [Permissions](http://swcarpentry.github.io/shell-extras/04-permissions/)


## Download data
Run this chunk to download the data for this challenge.

```{bash, download_data, engine.opts='-l'}
wget --no-verbose --no-check-certificate -O data-shell.zip https://swcarpentry.github.io/shell-novice/data/data-shell.zip
unzip -oq data-shell.zip
rm -f data-shell.zip
```

You should *NOT* add `data-shell/` to your git repository!  You might want to put `data-shell/` in the [.gitignore](https://gitlab.oit.duke.edu/IBIEM/IBIEM_2018_2019/blob/master/content/bootcamp/040_git_overview.md#ignoring-things) for this repo so that it doesn't show up in your Git pane.


## Navigating Files and Directories

### Getting Help
There are several ways to get information on a given program.  The internet can be useful, but it should be used with some caution, because the version you have installed might be different than the version that is documented.  The safest starting point is to use help information built into the computer you are using.  The easiest way is to use the help information built into the program itself by running the command with `-h` or `--help` flags (some programs both will work, some it is one or the other) for example:
```{bash get_help, error=TRUE}
cd --help
```

For the command `ls -lh` what effect does the `-lh` argument have (hint: single letter flags can often be combined , so `-lh` is equvalent to `-l -h`).  Show how you figure this out in the chunk below, then use words to give the answer below the chunk. It might also be a good idea to run `ls`, `ls -h`, `ls -l`, `ls -lh` to get a better feel for the effects of these flags.

```{bash}

```





## Working With Files and Directories
Run the chunk at the top of this challenge to create a directory named `data-shell` that contains a number of files and subdirectories.  

Now, make a new directory called `text_files` that is a subdirectory of `data-shell/data`, then move all of the files in `data-shell/data` that end in ".txt" to the `text_files` subdirectory, and list the contents of `data-shell/data` and `data-shell/data` `data-shell/data/text_files` to confirm that the files were moved.  
*Hint:* in your command you should not list each ".txt" file individually.

### Moving and Wildcards
```{bash}

```

### Creating Directories
In the `data-shell` directory make a new directory and subdirectories with the following structure, then use the `tree` command to confirm that it worked by printing out a tree that looks just like this one:

```
field_notes/
└── 2018
    ├── winter
    ├── spring
    ├── summer
    └── fall
```

```{bash}

```


## Bonus: Bash Variables

### Defining and Using Bash Variables
Bash is a full fledged programming language.  Like other programming languages, bash supports variables.  Here is an example of defining and using variables in bash.

```{bash variable_example}
FAVORITE_FRUIT="apples"

echo "I like ${FAVORITE_FRUIT}."
```

Notice that we define a variable in bash using just the variable name `FAVORITE_FRUIT`, but to use the varible we refer to it by the variable name prefixed with a dollar sign `$FAVORITE_FRUIT`, or the variable name enclosed in curly brackets, then prefixed with a dollar sign `${FAVORITE_FRUIT}`.


### Care and Feeding of Bash Variables
One difference between bash and most other programming languages is that, by default, it doesn't complain if you use a variable that you haven't defined, bash just assumes that the value of an undefined variable is an empty string

```{bash undefined_vars}
echo "I like drinking ${FAVORITE_DRINK}."
```

It can be dangerous to use undefined variables, so it is a good idea to set bash so that it *does* complain if you try to use an undefined variable, you can do this with the command `set -u`.
```{bash set_u, error=TRUE}
set -u
echo "I like drinking ${FAVORITE_DRINK}."
```

### Global Bash Variables in RMarkdown
One quirk of bash chunks in Rmarkdown documents is that each bash chunk has it's own environment, which is independent of the other chunks.  This means that variables defined in one bash chunk are not available in other chunks.  Remmeber we defined `$FAVORITE_FRUIT` at the top of this section, but the chunk below doesn't know anything about it!

```{bash independent_chunks, error=TRUE}
set -u
echo "I like ${FAVORITE_FRUIT}."
```

One hack that allows you to share bash variables between bash chunks is to set a bash variable in an *R chunk* using the R function `Sys.setenv`.  

Look up how to use `Sys.setenv`, then use the R chunk below to define the bash variable `$FAVORITE_DESSERT` so that the bash chunk below it runs correctly.

```{r}

```


```{bash test_Sys_setenv}
set -u
echo "My favorite dessert is ${FAVORITE_DESSERT}."
```




# Session Info
Always print `sessionInfo` for reproducibility!
```{r}

```

